Typesafe Coordinate Systems in High-Throughput Sequencing Applications

### 2022-02-23 manuscript/main.tex
added some manual page breaks to avoid breaking over some of the code excerpts

Instead, consider strongly and statically typed languages with generic and polymorphic features. Define a type class \texttt{Interval} of which there are four types, corresponding to ZBHO, ZBC, OBHO, and OBC coordinate systems, and for which there exists a function, \textit{to}, (or conversely, \textit{from}) that maps between constituent types. Note that the types may be freely interconverted, as there exists a bijective mapping between any two coordinate systems (Figure \ref{fig:transitiongraph}). While these types could all be defined individually, one might instead leverage polymorphic features of the type system, especially in systems where a type can be parameterized on a constant value. In C++, D, or similar languages this could be done by templating a single \texttt{Interval} class or struct on some enums defining basis and openness:

\begin{verbatim}
    Interval <CoordSystem::zbho> i (0, 100);    // C++
    auto i = Interval!CoordSystem.zbho(0, 100); // D
    let i = Interval::<ZeroBased, HalfOpen>::from_int(0, 100); // rust
\end{verbatim}

@@

As a concrete example, consider reading a coordinate range encompassing a gene or exon of interest from a GFF3 gene annotation file. In a typesafe coordinate system-enabled library such as \texttt{dhtslib}, the return value of any GFF3 reader will be of a type encoding its one-based closed (OBC) nature. Suppose then that the software must query a BAM file for alignments falling within the range just queried. In a typesafe coordinate system, the prior result can be passed directly to query function which is generic over the type of input coordinates and at compile time has logic injected to perform any necessary conversion of \textit{basis} and/or \textit{openness}. A pseudocode\footnote{Currently we don't have a good gene-based query method in the GFF3 module, but we should! and, the current iteration of Interval does not encode contig name/id, but we could potentially add this via an \texttt{Optional!} field} excerpt is shown below.\\

>>>>

Instead, consider strongly and statically typed languages with generic and polymorphic features. Define a type class \texttt{Interval} of which there are four types, corresponding to ZBHO, ZBC, OBHO, and OBC coordinate systems, and for which there exists a function, \textit{to}, (or conversely, \textit{from}) that maps between constituent types. Note that the types may be freely interconverted, as there exists a bijective mapping between any two coordinate systems (Figure \ref{fig:transitiongraph}). While these types could all be defined individually, one might instead leverage polymorphic features of the type system, especially in systems where a type can be parameterized on a constant value. 
\filbreak
In C++, D, or similar languages this could be done by templating a single \texttt{Interval} class or struct on some enums defining basis and openness:
\begin{verbatim}
    Interval <CoordSystem::zbho> i (0, 100);    // C++
    auto i = Interval!CoordSystem.zbho(0, 100); // D
    let i = Interval::<ZeroBased, HalfOpen>::from_int(0, 100); // rust
\end{verbatim}

@@

As a concrete example, consider reading a coordinate range encompassing a gene or exon of interest from a GFF3 gene annotation file. In a typesafe coordinate system-enabled library such as \texttt{dhtslib}, the return value of any GFF3 reader will be of a type encoding its one-based closed (OBC) nature. Suppose then that the software must query a BAM file for alignments falling within the range just queried. In a typesafe coordinate system, the prior result can be passed directly to query function which is generic over the type of input coordinates and at compile time has logic injected to perform any necessary conversion of \textit{basis} and/or \textit{openness}.

\filbreak
A pseudocode\footnote{Currently we don't have a good gene-based query method in the GFF3 module, but we should! and, the current iteration of Interval does not encode contig name/id, but we could potentially add this via an \texttt{Optional!} field} excerpt is shown below.\\
